### 2014-08-25 assets/paperdraft/parts/abstract.tex
splitted into several parts

% Abstract must be short enough to appear in this page together with title, authors, their addresses and keywords.
The use of multivariate classifiers has become commonplace in particle physics. 
Typically, a series of classifiers is trained rather than just one to enhance the performance; this is known as boosting.

In some applications of high energy physics (i.e., amplitude analyses) 
classifiers should not only optimize some integrated figure of merit, but produce a uniform selection efficiency in a space of selected physical variables.
Recently uBoost technique of boosting was proposed which addresses this issue.

In this paper we introduce several approaches to measure the uniformity of efficiency and propose new boosting techniques.