### 2014-09-09 assets/paperdraft/paper_draft.tex
Mike edits

\pdfoutput=1 % only if pdf/png/jpg images are used
\documentclass{JINST}

\title{New approaches for boosting to uniformity}

\author{
Aleksandar Bukva$^{\rm ADD}$, 
Vladimir Gligorov$^d$,
Alex Rogozhnikov$^{a,b}$\thanks{Corresponding author.}~ ,
Andrey Ustuzhanin$^b$ and
Mike Williams$^c$\\
\llap{$^a$}Lomonosov Moscow State University, Moscow, Russia\\
\llap{$^b$}Yandex LLC, Moscow, Russia\\
\llap{$^c$}Massachusetts Institute of Technology, Cambridge, MA, United States \\
\llap{$^d$}Organisation Europ\'eenne pour la Recherche Nucl\'eaire (CERN), Geneva, Switzerland  \\
E-mail: \email{alex.rogozhnikov@yandex.ru}}

\abstract{
\input{parts/abstract}
}

% AMS packages:
\usepackage{amsmath, amsthm, amsfonts, xspace}

% Theorems
%-----------------------------------------------------------------

\newtheorem{thm}{Theorem}[section]
\newtheorem{cor}[thm]{Corollary}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}
\theoremstyle{definition}
\newtheorem{defn}[thm]{Definition}
\theoremstyle{remark}
\newtheorem{rem}[thm]{Remark}

\def\RR{\mathbb{R}}
\def\ZZ{\mathbb{Z}}
\newcommand{\abs}[1]{\left\vert#1\right\vert}
\newcommand{\sgn}{\operatorname{sgn}}

%-----------------------------------------------------------------

\def \x {\ensuremath{\vec{x}}\xspace}
\def \y {\ensuremath{\vec{y}}\xspace}
\def\knn{\text{knn}}
\def\knni{\text{knn}(i)}

\begin{document}
\maketitle

\section{Introduction}

\input{parts/introduction}

%\section{Uniformity Measurement}
%\input{parts/metrics}
%\section{Approaches Proposed}
%\input{parts/approaches}

\section{Uniformity Boosting Methods}
\input{parts/methods}

\section{Example Analysis}

\input{parts/dataset}
\input{parts/plots}

\section{Timings}

\input{parts/timings}

\section{Summary}

\input{parts/summary}

\section{Source code}

%\textcolor{red}{xxx}

The link on the repository ans links to final notebooks with results in the article.

\acknowledgments

\begin{thebibliography}{9}

\bibitem{ref:lhcbhlt} R. Aaij {\em et al.} [LHCb Trigger Group], {\em The LHCb trigger and its performance}, JINST {\bf 8}, P04022 (2013). [arXiv:1211.3055]

\bibitem{ref:bbdt}  V. Gligorov and M. Williams, {\em Efficient, reliable and fast high-level triggering using a bonsai boosted decision tree}, JINST {\bf 8}, P02013 (2013). [arXiv:1210.6861]

\bibitem{ref:uboost} J. Stevens and M. Williams, {\em uBoost: A boosting method for producing uniform selection efficiencies from multivariate classifiers}, JINST {\bf 8}, P12013 (2013) [arXiv:1305.7248]

\end{thebibliography}

\appendix
\section{Other Flatness Loss Criteria}

ADD THEIL, ETC DESCRIPTIONS AND HOW TO IMPLEMENT IN THE ALGORITHM HERE.

\end{document}