### 2014-09-18 assets/paperdraft/parts/dataset.tex
First draft

\section{Datasets}
This study uses simulated event samples produced using the official LHCb simulation framework.
The software used for the generation of the events is described in LHCb publications as follows :

\begin{quote}
In the simulation, $pp$ collisions are generated using
PYTHIA~\cite{Sjostrand:2006za} 
with a specific LHCb configuration~\cite{LHCb-PROC-2010-056}.  Decays of hadronic particles
are described by EvtGen~\cite{Lange:2001uf}, in which final state
radiation is generated using PHOTOS~\cite{Golonka:2005pn}. The
interaction of the generated particles with the detector and its
response are implemented using the GEANT
toolkit~\cite{Allison:2006ve, Agostinelli:2002hh} as described in
Ref.~\cite{LHCb-PROC-2011-006}.
\end{quote}

All simulated event samples are generated inside the LHCb detector acceptance.
The signal used in this analysis consists of $D_s^\pm\to\pi^+\pi^-\pi^\pm$ decays, simulated
using the $\textrm{D}\_\textrm{DALITZ}$ model of EvtGen to simulate the intermediate resonances which contribute to the
three pion final state. The background candidates are three pion combinations reconstructed in
simulated samples of $c\bar{c}$ and $b\bar{b}$ events, where the charm and bottom quark decays are
inclusively modelled by EvtGen. The simulated events contain ``truth'' information which identifies them as
signal or background, and which identifies the physical origin of the three pion combinations reconstructed
in the $c\bar{c}$ and $b\bar{b}$ simulated samples.