### 2014-09-18 assets/paperdraft/paper_draft.tex
Add references

\pdfoutput=1 % only if pdf/png/jpg images are used
\documentclass{JINST}

\title{New Approaches to Uniform Boosting}

\author{
Aleksandar Bukva$^e$, 
Vladimir Gligorov$^d$,
Alex Rogozhnikov$^{a,b}$\thanks{Corresponding author.}~ ,
Andrey Ustuzhanin$^b$ and
Mike Williams$^c$\\
\llap{$^a$}Lomonosov Moscow State University, Moscow, Russia\\
\llap{$^b$}Yandex LLC, Moscow, Russia\\
\llap{$^c$}Massachusetts Institute of Technology, Cambridge, MA, United States \\
\llap{$^d$}Organisation Europ\'eenne pour la Recherche Nucl\'eaire (CERN), Geneva, Switzerland  \\
\llap{$^e$}Faculty of Physics, Belgrade, Serbia \\
E-mail: \email{alex.rogozhnikov@yandex.ru}}

\abstract{
\input{parts/abstract}
}

% AMS packages:
\usepackage{amsmath, amsthm, amsfonts}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{float}

% Theorems
%-----------------------------------------------------------------

\newtheorem{thm}{Theorem}[section]
\newtheorem{cor}[thm]{Corollary}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}
\theoremstyle{definition}
\newtheorem{defn}[thm]{Definition}
\theoremstyle{remark}
\newtheorem{rem}[thm]{Remark}

\def\RR{\mathbb{R}}
\def\ZZ{\mathbb{Z}}
\newcommand{\abs}[1]{\left\vert#1\right\vert}
\newcommand{\sgn}{\operatorname{sgn}}

%-----------------------------------------------------------------

\begin{document}
\maketitle

\section{Introduction}

\input{parts/introduction}

\section{Uniformity Measurement}

\input{parts/metrics}

\section{Approaches Proposed}

\input{parts/approaches}

\section{Dataset}

\input{parts/dataset}

\section{Plots}

\input{parts/plots}

\section{Timings}

\input{parts/timings}

\section{Summary}

\input{parts/summary}

\section{Source code}

The link on the repository ans links to final notebooks with results in the article.

\acknowledgments

These results were obtained using events generated with the official LHCb simulation, and we are grateful to the LHCb collaboration for this privilege. 
We particularly acknowledge the work of the LHCb Simulation and Core Computing teams in tuning the simulation software and managing the productions of simulated events.

\begin{thebibliography}{9}

\bibitem{Sjostrand:2006za}
{"Sj\"{o}strand, Torbj\"{o}rn and Mrenna, Stephen and Skands, Peter"},
\emph{PYTHIA 6.4 physics and manual},
\href{http://dx.doi.org/10.1088/1126-6708/2006/05/026},
{\emph{JHEP} {\bf 05} (2006) 026},
[\hepph{0603175}].

\bibitem{LHCb-PROC-2010-056}
{"Belyaev, I. and others"},
\emph{Handling of the generation of primary events in GAUSS, the LHCb simulation framework},
\href{http://dx.doi.org/10.1109/NSSMIC.2010.5873949},
{\emph{NSS/MIC} {\bf IEEE} (2010) 1155}.

\bibitem{Lange:2001uf}
{"Lange, D. J."},
\emph{The EvtGen particle decay simulation package},
\href{http://dx.doi.org/10.1016/S0168-9002(01)00089-4},
{\emph{NIM} {\bf A462} (2001) 152-155}.

\bibitem{Golonka:2005pn}
{"Golonka, Piotr and Was, Zbigniew"},
\emph{PHOTOS Monte Carlo: a precision tool for QED corrections in $Z$ and $W$ decays},
\href{http://dx.doi.org/10.1140/epjc/s2005-02396-4},
{\emph{EPJ} {\bf C45} (2006) 97-107},
[\hepph{0506026}].

\bibitem{Allison:2006ve}
{"Allison, John and Amako, K. and Apostolakis, J. and Araujo, H. and Dubois, P.A. and others"},
\emph{Geant4 developments and applications},
\href{http://dx.doi.org/10.1109/TNS.2006.869826},
{\emph{IEEE TNS} {\bf 53} (2006) 270}.

\bibitem{Agostinelli:2002hh}
{"Agostinelli, S. and others"},
\emph{Geant4: a simulation toolkit},
\href{http://dx.doi.org/10.1016/S0168-9002(03)01368-8},
{\emph{NIM} {\bf A506} (2003) 250}.

\bibitem{LHCb-PROC-2011-006}
{"Clemencic, M and others"},
\emph{The LHCb simulation application, GAUSS : design, evolution and experience},
\href{http://dx.doi.org/10.1088/1742-6596/331/3/032023},
{\emph{J. Phys. Conf. Ser.} {\bf 331} (2011) 032023}.

% The examples of items:
% \bibitem{bib1}
% Authors,
% \emph{Title},
% \emph{J. Ref.} \textbf{vol} (year) page.

% \bibitem{bib2}
% A. Pietropaolo et al., \emph{DINS measurements on VESUVIO in the
%     resonance detector configuration: proton mean kinetic energy in
%     water}, \jinst{1}{2006}{P04001}.

% \bibitem{bib3}
% A.I. Harris,
% \emph{Spectroscopy with multichannel correlation radiometers},
% \href{http://dx.doi.org/10.1063/1.1898643}
% {\emph{Rev.\ Sci.\ Instrum.} {\bf 76} (2005) 054503}
% [\astroph{0504449}].

% \bibitem{bib4}
% G.F. Knoll, \emph{Radiation detection and measurements}, John Wiley
%     and Sons, Inc., New York 2000.

% \bibitem{bib5}
% V. Dangendorf, \emph{Time-resolved fast-neutron imaging with a
% pulse-counting image intensifier}, in proceedings of
% \emph{International workshop on fast neutron detectors and
% applications}, April, 3--6, 2006 University of Cape Town, South Africa
% \pos{PoS(FNDA2006)008}.

\end{thebibliography}

\end{document}