### 2014-09-20 assets/paperdraft/paper_draft.tex

\pdfoutput=1 % only if pdf/png/jpg images are used
\documentclass{JINST}

\title{New approaches for boosting to uniformity}

\author{
Aleksandar Bukva$^e$, 
Vladimir Gligorov$^d$,
Alex Rogozhnikov$^{a,b}$\thanks{Corresponding author.}~,\newline
Andrey Ustyuzhanin$^{b,f}$ and
Mike Williams$^c$\\
\llap{$^a$}Lomonosov Moscow State University, Moscow, Russia\\
\llap{$^b$}Yandex, Moscow, Russia\\
\llap{$^c$}Massachusetts Institute of Technology, Cambridge, MA, United States \\
\llap{$^d$}Organisation Europ\'eenne pour la Recherche Nucl\'eaire (CERN), Geneva, Switzerland  \\
\llap{$^e$}Faculty of Physics, Belgrade, Serbia \\
\llap{$^f$}Moscow Institute of Physics and Technology, Moscow, Russia\\
E-mail: \email{alex.rogozhnikov@yandex.ru}}

\abstract{
\input{parts/abstract}
}

% AMS packages:

\usepackage{amsmath, amsthm, amsfonts, xspace}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{float}

% Theorems
%-----------------------------------------------------------------

\newtheorem{thm}{Theorem}[section]
\newtheorem{cor}[thm]{Corollary}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}
\theoremstyle{definition}
\newtheorem{defn}[thm]{Definition}
\theoremstyle{remark}
\newtheorem{rem}[thm]{Remark}

\def\RR{\mathbb{R}}
\def\ZZ{\mathbb{Z}}
\newcommand{\abs}[1]{\left\vert#1\right\vert}
\newcommand{\sgn}{\operatorname{sgn}}

%-----------------------------------------------------------------

\def\x {\ensuremath{\vec{x}}\xspace}
\def\y {\ensuremath{\vec{y}}\xspace}
\def\knn{\text{knn}}
\def\knni{\text{knn}(i)}

\begin{document}
\maketitle

\section{Introduction}

\input{parts/introduction}

\section{Uniformity Boosting Methods}
\input{ugb}

\section{Example Analysis}

\input{parts/dataset}
\input{parts/plots}

\section{Timings}

\input{parts/timings}

\section{Summary}

\input{parts/summary}

\section{Source code}

%\textcolor{red}{xxx}

The link on the repository ans links to final notebooks with results in the article.

\acknowledgments

These results were obtained using events generated with the official LHCb simulation, and we are grateful to the LHCb collaboration for this privilege. 
We particularly acknowledge the work of the LHCb Simulation and Core Computing teams in tuning the simulation software and managing the productions of simulated events.

\begin{thebibliography}{9}

\bibitem{ref:lhcbhlt} 
R. Aaij {\em et al.} [LHCb Trigger Group], 
{\em The LHCb trigger and its performance}, 
JINST {\bf 8}, P04022 (2013). 
[arXiv:1211.3055]

\bibitem{ref:bbdt}  
V. Gligorov and M. Williams, 
{\em Efficient, reliable and fast high-level triggering using a bonsai boosted decision tree}, 
JINST {\bf 8}, P02013 (2013). 
[arXiv:1210.6861]

\bibitem{ref:uboost} 
J. Stevens and M. Williams,
{\em uBoost: A boosting method for producing uniform selection efficiencies from multivariate classifiers}, 
JINST {\bf 8}, P12013 (2013). 
[arXiv:1305.7248]

\bibitem{Sjostrand:2006za}
{"Sj\"{o}strand, Torbj\"{o}rn and Mrenna, Stephen and Skands, Peter"},
\emph{PYTHIA 6.4 physics and manual},
\href{http://dx.doi.org/10.1088/1126-6708/2006/05/026},
{\emph{JHEP} {\bf 05} (2006) 026}.
[\hepph{0603175}]

\bibitem{LHCb-PROC-2010-056}
{"Belyaev, I. and others"},
\emph{Handling of the generation of primary events in GAUSS, the LHCb simulation framework},
\href{http://dx.doi.org/10.1109/NSSMIC.2010.5873949},
{\emph{NSS/MIC} {\bf IEEE} (2010) 1155}.

\bibitem{Lange:2001uf}
{"Lange, D. J."},
\emph{The EvtGen particle decay simulation package},
\href{http://dx.doi.org/10.1016/S0168-9002(01)00089-4},
{\emph{NIM} {\bf A462} (2001) 152-155}.

\bibitem{Golonka:2005pn}
{"Golonka, Piotr and Was, Zbigniew"},
\emph{PHOTOS Monte Carlo: a precision tool for QED corrections in $Z$ and $W$ decays},
\href{http://dx.doi.org/10.1140/epjc/s2005-02396-4},
{\emph{EPJ} {\bf C45} (2006) 97-107}.
[\hepph{0506026}]

\bibitem{Allison:2006ve}
{"Allison, John and Amako, K. and Apostolakis, J. and Araujo, H. and Dubois, P.A. and others"},
\emph{Geant4 developments and applications},
\href{http://dx.doi.org/10.1109/TNS.2006.869826},
{\emph{IEEE TNS} {\bf 53} (2006) 270}.

\bibitem{Agostinelli:2002hh}
{"Agostinelli, S. and others"},
\emph{Geant4: a simulation toolkit},
\href{http://dx.doi.org/10.1016/S0168-9002(03)01368-8},
{\emph{NIM} {\bf A506} (2003) 250}.

\bibitem{LHCb-PROC-2011-006}
{"Clemencic, M and others"},
\emph{The LHCb simulation application, GAUSS : design, evolution and experience},
\href{http://dx.doi.org/10.1088/1742-6596/331/3/032023},
{\emph{J. Phys. Conf. Ser.} {\bf 331} (2011) 032023}.

% The examples of items:
% \bibitem{bib1}
% Authors,
% \emph{Title},
% \emph{J. Ref.} \textbf{vol} (year) page.

% \bibitem{bib2}
% A. Pietropaolo et al., \emph{DINS measurements on VESUVIO in the
%     resonance detector configuration: proton mean kinetic energy in
%     water}, \jinst{1}{2006}{P04001}.

% \bibitem{bib3}
% A.I. Harris,
% \emph{Spectroscopy with multichannel correlation radiometers},
% \href{http://dx.doi.org/10.1063/1.1898643}
% {\emph{Rev.\ Sci.\ Instrum.} {\bf 76} (2005) 054503}
% [\astroph{0504449}].

% \bibitem{bib4}
% G.F. Knoll, \emph{Radiation detection and measurements}, John Wiley
%     and Sons, Inc., New York 2000.

% \bibitem{bib5}
% V. Dangendorf, \emph{Time-resolved fast-neutron imaging with a
% pulse-counting image intensifier}, in proceedings of
% \emph{International workshop on fast neutron detectors and
% applications}, April, 3--6, 2006 University of Cape Town, South Africa
% \pos{PoS(FNDA2006)008}.

\end{thebibliography}

\appendix
\section{Other Flatness Loss Criteria}

Clean all this up...

%\section{Uniformity Measurement}
\input{parts/metrics}
%\section{Approaches Proposed}
\input{parts/approaches}

\end{document}