### 2014-10-07 assets/paperdraft/paper_draft.tex
Mike edits

\pdfoutput=1 % only if pdf/png/jpg images are used
\documentclass{JINST}

\title{New approaches for boosting to uniformity}

\author{
Alex Rogozhnikov$^{a,b}$\thanks{Corresponding author.}~,
Aleksandar Bukva$^c$, 
Vladimir Gligorov$^d$,
Andrey Ustyuzhanin$^{b,e,f}$ and
Mike Williams$^g$\\
\llap{$^a$}Lomonosov Moscow State University, Moscow, Russia\\
\llap{$^b$}Yandex, Moscow, Russia\\
\llap{$^c$}Faculty of Physics, Belgrade, Serbia \\
\llap{$^d$}Organisation Europ\'eenne pour la Recherche Nucl\'eaire (CERN), Geneva, Switzerland  \\
\llap{$^e$}Moscow Institute of Physics and Technology, Moscow, Russia\\
\llap{$^f$}Imperial College, London, UK\\
\llap{$^g$}Massachusetts Institute of Technology, Cambridge, MA, United States \\
E-mail: \email{alex.rogozhnikov@yandex.ru}}

\abstract{
\input{parts/abstract}
}

% AMS packages:

\usepackage{amsmath, amsthm, amsfonts, xspace}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{float}

% Theorems
%-----------------------------------------------------------------

\newtheorem{thm}{Theorem}[section]
\newtheorem{cor}[thm]{Corollary}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}
\theoremstyle{definition}
\newtheorem{defn}[thm]{Definition}
\theoremstyle{remark}
\newtheorem{rem}[thm]{Remark}

\def\RR{\mathbb{R}}
\def\ZZ{\mathbb{Z}}
\newcommand{\abs}[1]{\left\vert#1\right\vert}
\newcommand{\sgn}{\operatorname{sgn}}

%-----------------------------------------------------------------

\def\x {\ensuremath{\vec{x}}\xspace}
\def\y {\ensuremath{\vec{y}}\xspace}
\def\knn{\text{knn}}
\def\knni{\text{knn}(i)}

\begin{document}
\maketitle

\section{Introduction}

\input{parts/introduction}

\section{Uniformity Boosting Methods}
\input{ugb}

\section{Example Analysis}

\input{parts/dataset}
\input{parts/plots}

\section{CPU Resources}

\input{parts/timings}

\section{Summary}

\input{parts/summary}

\section{Source code}

UPDATE THIS WITH CODE LOCATION.
%The link on the repository ans links to final notebooks with results in the article.

\acknowledgments

These results were obtained using events generated with the official LHCb simulation, and we are grateful to the LHCb collaboration for this privilege. 
We particularly acknowledge the work of the LHCb Simulation and Core Computing teams in tuning the simulation software and managing the productions of simulated events.
MW is supported by NSF grant PHY-1306550.

\begin{thebibliography}{9}

\bibitem{ref:lhcbhlt} 
R. Aaij {\em et al.} [LHCb Trigger Group], 
{\em The LHCb trigger and its performance}, 
JINST {\bf 8}, P04022 (2013). 
[arXiv:1211.3055]

\bibitem{ref:bbdt}  
V. Gligorov and M. Williams, 
{\em Efficient, reliable and fast high-level triggering using a bonsai boosted decision tree}, 
JINST {\bf 8}, P02013 (2013). 
[arXiv:1210.6861]

\bibitem{ref:FS1997}
{Y. Freund, and R. Schapire},
\emph{A Decision-Theoretic Generalization of On-Line Learning and an Application to Boosting}, 
{\emph{Journal of Computer and System Sciences} {\bf 55(1)} (August 1997) 119-139}.

\bibitem{ref:uboost} 
J. Stevens and M. Williams,
{\em uBoost: A boosting method for producing uniform selection efficiencies from multivariate classifiers}, 
JINST {\bf 8}, P12013 (2013). 
[arXiv:1305.7248]

\bibitem{ref:F1999}
{Jerome H. Friedman},
\emph{Greedy Function Approximation: A Gradient Boosting Machine},
\href{http://statweb.stanford.edu/~jhf/ftp/trebst.pdf}
{\emph{Annals of Statistics} {\bf 29} (2000) 1189-1232}.

\bibitem{Sjostrand:2006za}
T.~Sj\"{o}strand, S.~Mrenna and P.~Skands,
\emph{PYTHIA 6.4 physics and manual},
\href{http://dx.doi.org/10.1088/1126-6708/2006/05/026}
{\emph{JHEP} {\bf 05} (2006) 026}.
[\hepph{0603175}]

\bibitem{LHCb-PROC-2010-056}
I.~Belyaev, I. {\em et al.},
\emph{Handling of the generation of primary events in GAUSS, the LHCb simulation framework},
\href{http://dx.doi.org/10.1109/NSSMIC.2010.5873949}
{\emph{NSS/MIC} {\bf IEEE} (2010) 1155}.

\bibitem{Lange:2001uf}
D.J.~Lange,
\emph{The EvtGen particle decay simulation package},
\href{http://dx.doi.org/10.1016/S0168-9002(01)00089-4}
{\emph{NIM} {\bf A462} (2001) 152-155}.

\bibitem{Golonka:2005pn}
P.~Golonka and Z.~Was,
\emph{PHOTOS Monte Carlo: a precision tool for QED corrections in $Z$ and $W$ decays},
\href{http://dx.doi.org/10.1140/epjc/s2005-02396-4}
{\emph{EPJ} {\bf C45} (2006) 97-107}.
[\hepph{0506026}]

\bibitem{Allison:2006ve}
J.~Allison {\em et al.},
\emph{Geant4 developments and applications},
\href{http://dx.doi.org/10.1109/TNS.2006.869826}
{\emph{IEEE TNS} {\bf 53} (2006) 270}.

\bibitem{Agostinelli:2002hh}
S.~Agostinelli {\em et al.},
\emph{Geant4: a simulation toolkit},
\href{http://dx.doi.org/10.1016/S0168-9002(03)01368-8}
{\emph{NIM} {\bf A506} (2003) 250}.

\bibitem{LHCb-PROC-2011-006}
M~Clemencic {\em et al.},
\emph{The LHCb simulation application, GAUSS : design, evolution and experience},
\href{http://dx.doi.org/10.1088/1742-6596/331/3/032023}
{\emph{J. Phys. Conf. Ser.} {\bf 331} (2011) 032023}.

% The examples of items:
% \bibitem{bib1}
% Authors,
% \emph{Title},
% \emph{J. Ref.} \textbf{vol} (year) page.

% \bibitem{bib2}
% A. Pietropaolo et al., \emph{DINS measurements on VESUVIO in the
%     resonance detector configuration: proton mean kinetic energy in
%     water}, \jinst{1}{2006}{P04001}.

% \bibitem{bib3}
% A.I. Harris,
% \emph{Spectroscopy with multichannel correlation radiometers},
% \href{http://dx.doi.org/10.1063/1.1898643}
% {\emph{Rev.\ Sci.\ Instrum.} {\bf 76} (2005) 054503}
% [\astroph{0504449}].

% \bibitem{bib4}
% G.F. Knoll, \emph{Radiation detection and measurements}, John Wiley
%     and Sons, Inc., New York 2000.

% \bibitem{bib5}
% V. Dangendorf, \emph{Time-resolved fast-neutron imaging with a
% pulse-counting image intensifier}, in proceedings of
% \emph{International workshop on fast neutron detectors and
% applications}, April, 3--6, 2006 University of Cape Town, South Africa
% \pos{PoS(FNDA2006)008}.

\end{thebibliography}

\appendix
%\section{Uniformity Measurement}
\input{parts/metrics}
%\section{Approaches Proposed}
%\input{parts/approaches}

\end{document}