### 2014-10-07 assets/paperdraft/parts/timings.tex

One drawback of the uBoost technique is that it has a high degree of computational complexity: 
while AdaBoost trains $M$ trees (a user-defined number), uBoost builds $100 \times M$ trees.  
The algorithms presented in this paper only build $M$ trees; however, the boosting involves some more complicated algorithms.  Training each of the $M$ trees scales as follows for $N$ training events:
\begin{itemize}
       \item kNNAdaBoost: $O(k \times N)$;
	\item uGBkNNknn: $O(k \times N)$ for $A_{\rm knn}$, and  
	$O( \text{\#nonzero elements in the matrix})$ for arbitrary matrix $A$;
	\item uGBFL(bin): $O(N \ln N)$;
	\item uGBFL(kNN): $O(N \ln N + N k \ln k) $.
\end{itemize}
In the example analysis studied in this paper, we find that the training time for these new algorithms is within a factor of two the same as AdaBoost.  The CPU-resource usage of these new algorithms is not prohibitive.